Visualization tool for comparison of a single amino acid change in protein simulation

### 2017-08-03 paper/paper.tex
paper

% DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.
%
\end{abstract}

@@

\subsection{Related work}
%
The work in~\cite{likhachev2016available} reviews some of the available instruments when analyzing molecular dynamics trajectories, including visualization tools. OVITO~\cite{stukowski2009visualization} is an open visualization tool for atomistic simulation data. TRAVIS~\cite{brehm2012travis} is a free analyzer and visualizer for Monte Carlo and molecular dynamics trajectories. The study in~\cite{garg2010visualization} presents a visualization tool for many tasks of molecular dynamics simulation. QwikMD~\cite{ribeiro2016qwikmd} is a toolkit that includes tools for analyzing simulations.

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figs/grid.png}
\caption{Overview of the used grid.} 
\label{fig:grid}
\end{figure}
\paragraph*{Item highlighting} When hoving the mouse over a line in the grid, that line is highlited. This operation also reflects on the other charts.
\paragraph*{Item selection} When right-clicking on a line of the grid with the mouse that line is selected and the item is fixed on the charts. The selected line is becomes darker on the parallel coordinates chart.

@@

The objective of the radar chart~\cite{saary2008radar}~\cite{mosley1999benchmarking}, showed in Fig.~\ref{fig:radar}, is to visualize and compare specif items of the data, i.e., single simulations. The hypothesis is that overlaping the shapes of the lines from the parallel coordinates in a radial way makes it easier com see the similarities and differences between a small number of items, while the global view is better represented by the parallel coordinates chart since for many shapes the radar chart becomes convoluted.

@@

All elements of this visualization tool were implemented using JavaScript, HTML and CSS, with the visualization library D3.js. The parallel coordinates chart uses the visual toolkit for multidimensional detectives Parallel Coordinates (0.7.0) (\url{http://syntagmatic.github.io/parallel-coordinates}).

The radar chart was build from the implementations of Nadieh Bremer (\url{http://visualcinnamon.com/2015/10/different-look-d3-radar-chart.html}) and Micah Stubbs (\url{http://bl.ocks.org/micahstubbs/a772306d6fd49874ec92}).

As previously mentioned, the Kappa-conotoxin PVIIA was chosen as example for testing the visualization tool. We know from the start that the simulation containing the asparagine ($N$) in the fifth position of its amino acid chain represents the protein in its natural state. Also, the change of this asparagine by an alanine is reported to reduce 100\% of the proteins toxicity. When comparing the simulations with asparagine and alanine, thus, it would be expected to see some similarity between then, since the mutation don't destroy the protein, but also enough differences to justify the loss of function.

>>>>

% DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.
%

Molecular dynamics simulation is a powerful tool for understanding the behavior of biomolecules such as proteins. When comparing the trajectory convergence of simulations, however, the researcher can face difficulties inspecting and selecting the multidimensional data. In this paper it is proposed a new visualization tool that combines parallel coordinates, radar chart and grid and the ability to update simulation data over time. The tool is tested with simulations of mutations of the Kappa-conotoxin PVIIA.

\end{abstract}

@@

\subsection{Related work}
%
The work of Likhachev et al., 2016~\cite{likhachev2016available} reviews some of the available instruments when analyzing molecular dynamics trajectories, including visualization tools. OVITO~\cite{stukowski2009visualization} is an open visualization tool for atomistic simulation data. TRAVIS~\cite{brehm2012travis} is a free analyzer and visualizer for Monte Carlo and molecular dynamics trajectories. The study of Garg, 2010~\cite{garg2010visualization} presents a visualization tool for many tasks of molecular dynamics simulation. QwikMD~\cite{ribeiro2016qwikmd} is a toolkit that includes tools for analyzing simulations.

@@

\begin{figure}
\includegraphics[width=0.9\linewidth]{figs/grid.png}
\caption{Overview of the used grid.} 
\label{fig:grid}
\end{figure}
\paragraph*{Item highlighting} When hoving the mouse over a line in the grid, that line is highlited. This operation also reflects on the other charts.
\paragraph*{Item selection} When right-clicking on a line of the grid with the mouse that line is selected and the item is fixed on the charts. The selected line becomes darker on the parallel coordinates chart.

@@

The objective of the radar chart~\cite{saary2008radar}~\cite{mosley1999benchmarking}, showed in Fig.~\ref{fig:radar}, is to visualize and compare specific items of the data, i.e., single simulations. The hypothesis is that overlaping the shapes of the lines from the parallel coordinates in a radial way makes it easier to see the similarities and differences between a small number of items, while the global view is better represented by the parallel coordinates chart since for many shapes the radar chart becomes convoluted.

@@

All elements of this visualization tool were implemented using JavaScript, HTML and CSS, with the visualization library D3.js. The parallel coordinates chart uses the visual toolkit for multidimensional detectives Parallel Coordinates (0.7.0) \footnote{\url{http://syntagmatic.github.io/parallel-coordinates}}. The radar chart was build from the implementations of Nadieh Bremer \footnote{\url{http://visualcinnamon.com/2015/10/different-look-d3-radar-chart.html}} and Micah Stubbs \footnote{\url{http://bl.ocks.org/micahstubbs/a772306d6fd49874ec92}}.

As previously mentioned, the Kappa-conotoxin PVIIA was chosen as example for testing the visualization tool. We know from the start that the simulation containing the asparagine ($N$) in the fifth position of its amino acid chain represents the protein in its natural state. Also, the change of this asparagine by an alanine is reported to reduce 100\% of the proteins toxicity. When comparing the simulations with asparagine and alanine, thus, it would be expected to see some similarity between them, since the mutation does not destroy the protein, but also enough differences to justify the loss of function.
